\begin{itemize}[leftmargin=*]
    \item
        \textbf{Apache}
        A web-server
    \item
        \textbf{AWS (Amazon Web Services)}
        Various services run by Amazon for helping to create web-based infrastructure
    \item
        \textbf{Class}:
        An abstract representation of an object instance
    \item
        \textbf{Client-Side}:
        Code that runs on the user's computer. As opposed to server-side, which runs on the server.
    \item
        \textbf{Composer}
        PHP's package management system
    \item
        \textbf{Controller}:
        A piece of code that is run for a specific route, whose job it is to get/update the relevant data and return a response to the user.
    \item
        \textbf{CORS (Cross-Origin Resource Sharing)}:
        A safety feature that permits browsers to make requests to APIs on different domains.
    \item
        \textbf{Dependency Injection}
        Rather than hard-coding a dependency on a specific class, taking advantage of polymorphism and passing in a class that implements a specific interface.
    \item
        \textbf{Foreign Key}:
        A relationship between two tables in MySQL that is enforced by the database.
    \item
        \textbf{Instance}:
        An object is an instance of a specific class with its own set of properties
    \item
        \textbf{Many-to-Many}:
        A relationship between two tables of a database where the items of table A can be linked to many items from table B, and the items from table B can be linked to many items in table A.
    \item
        \textbf{Middleware}:
        A piece of code that runs before/after the controller code to change the request/response in some way.
    \item
        \textbf{Namespace}:
        A set of classes where each class has a unique name. We can have two classes with the same name as long as they are in different namespaces.
    \item
        \textbf{One-to-Many}:
        A relationship between two tables of a database where the items of table A can be linked to many items from table B, but items from table B can only be linked to one item in table A.
    \item
        \textbf{ORM (Object Relational Mapper)}:
        Allows us to access data from a database using standard objects.
    \item
        \textbf{Pivot Table}:
        A table that stores a many-to-many relationship.
    \item
        \textbf{Provisioning}:
        Setting up a server/virtual machine so that it's ready to be used.
    \item
        \textbf{Relational Database Management Systems}:
        A database made up of tables (that represent one type of thing) and the relationships between items in those tables.
    \item
        \textbf{Resource}:
        Allows us to control the JSON output of our API.
    \item
        \textbf{SSH (Secure SHell)}:
        A way to securely communicate with a remote machine over an insecure network
    \item
        \textbf{Static}:
        A property or method that belongs to a class rather than an object instance
    \item
        \textbf{Virtual Server}
        A server that runs as if you have complete control over a physical machine, but is actually one of many virtual machines on a single (powerful) piece of hardware
\end{itemize}
